---
title: "R Notebook"
output: md_document
---

### Load the libraries
```{r}
library(tidyverse)
library(arrow)
library(tidyverse)
library(lubridate)
library(gender)
library(igraph)
library(dplyr)
```

### Load the data - this is filtered & merged (with transaction) data
```{r}
applications <- read.csv("/Users/kaz/DataspellProjects/Org-Analytics/Final Project/Data/closest_dock_events.csv")
```


### Experience Variable
This variable is a measure of the number of years of experience of the applicant. It is calculated by subtracting the year of the "filing" from the year of the first publication of the applicant.
```{r}
# distribution of experience
ggplot(applications, aes(x = experience)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 100) +
  labs(title = "Distribution of Experience",
       x = "Experience (years)",
       y = "Frequency")
```


**Mean Proc Time by workgroup and show distribution of means**
```{r}
wg_means <- applications %>%
  group_by(workgroup) %>%
  summarise(mean_proc_time=mean(processing_time, na.rm = TRUE),
            n = n()) %>% arrange(desc(mean_proc_time))

# sort it
ggplot(wg_means, aes(x = reorder(workgroup, mean_proc_time), y = mean_proc_time)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Mean Processing Time by Workgroup",
       x = "Workgroup",
       y = "Mean Processing Time (days)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



**Application Status is limited to 150 and 161**
```{r}
applications %>%
  group_by(gender) %>%
  summarise(mean_app_proc_time=mean(processing_time, na.rm = TRUE),
            n = n()) %>% arrange(desc(n))


applications %>%
  group_by(race) %>%
  summarise(mean_app_proc_time=mean(processing_time, na.rm = TRUE),
            n = n()) %>% arrange(desc(n))
```


| Code | Status                                               | Frequency | Percent |
|------|------------------------------------------------------|-----------|---------|
| 150  | Patented Case                                        | 4,288,187 | 46.5%   |
| 161  | Abandoned -- Failure to Respond to an Office Action  | 1,107,772 | 12.0%   |
| 250  | Patent Expired Due to NonPayment of Maintenance Fees Under 37 CFR 1.362 | 919,203 | 10.0%   |
| 159  | Provisional Application Expired                     | 848,536   | 9.2%    |
| 30   | Docketed New Case - Ready for Examination           | 350,347   | 3.8%    |
| 218  | RO PROCESSING COMPLETED-PLACED IN STORAGE           | 284,700   | 3.1%    |
| 566  | PCT - International Search Report Mailed to IB      | 234,837   | 2.5%    |
| 41   | Non Final Action Mailed                              | 168,103   | 1.8%    |
| 166  | Abandoned -- File-Wrapper-Continuation Parent Application | 165,589 | 1.8%    |
| 19   | Application Undergoing Preexam Processing            | 126,741   | 1.4%    |
| N/A  | Other Codes                                          | 709,623   | 7.7%    |
|      | Missing                                              | 27,532    | 0.3%    |
|      | Total                                                | 9,231,170 | 100.0%  |




#### Sequence diff vs processing_time
```{r}
# Generate scatter plot of sequence_diff vs processing_time
ggplot(applications, aes(x = sequence_diff, y = processing_time)) +
  geom_point(alpha = 0.5) +  # alpha for transparency if many points overlap
  labs(x = "Sequence Difference", y = "Processing Time",
       title = "Scatter Plot of Sequence Difference vs Processing Time") +
  theme_minimal()  # A clean theme for the plot
```



```{r}

ggplot(applications, aes(x = out_degree, y = processing_time)) +
  geom_point(alpha = 0.5) +  # alpha for transparency if many points overlap
  labs(x = "out_degree", y = "Processing Time",
       title = "Scatter Plot of out_deg vs Processing Time") +
  theme_minimal()  # A clean theme for the plot
```


```{r}

ggplot(applications, aes(x = out_degree, y = processing_time)) +
  geom_point(alpha = 0.5) +  # alpha for transparency if many points overlap
  labs(x = "in_degree", y = "Processing Time",
       title = "Scatter Plot of out_deg vs Processing Time") +
  theme_minimal()  # A clean theme for the plot
```

```{r}

ggplot(applications, aes(x = betweenness, y = processing_time)) +
  geom_point(alpha = 0.5) +  # alpha for transparency if many points overlap
  labs(x = "betweenness", y = "Processing Time",
       title = "Scatter Plot of between vs Processing Time") +
  theme_minimal()  # A clean theme for the plot
```


```{r}

ggplot(applications, aes(x = advice_count_final)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 100) +
  labs(title = "Distribution of Experience",
       x = "advice_counts",
       y = "Frequency")
```



