---
title: "R Notebook"
output: md_document
---

### Load the libraries
```{r}
library(tidyverse)
library(arrow)
library(tidyverse)
library(lubridate)
library(gender)
library(igraph)
library(dplyr)
```

### Load the data - this is filtered & merged (with transaction) data
```{r}
applications <- read.csv("/Users/kaz/Desktop/Final Project_org/Data/closest_dock_events.csv")
```


### Experience Variable
This variable is a measure of the number of years of experience of the applicant. It is calculated by subtracting the year of the "filing" from the year of the first publication of the applicant.
```{r}
# distribution of experience
ggplot(applications, aes(x = experience)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 100) +
  labs(title = "Distribution of Experience",
       x = "Experience (days)",
       y = "Frequency")
```

```{r}
# experience by race gender and race
# distinct examiner_id and keep all columns
applications_unique <- applications %>% distinct(application_number, .keep_all = TRUE)

library(RColorBrewer)

ggplot(applications_unique, aes(x = gender, y = experience, fill = race)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Paired") +  # You can choose other palettes like "Dark2", "Set3", etc.
  labs(title = "Distribution of Experience by Gender and Race",
       x = "Gender",
       y = "Experience") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Adjust text angle if needed

```


```{r}
# create experience_year variable - floor the experience to years
applications$experience_years <- floor(applications$experience / 365)
```


### how does network feature vary by seniority (experience)
```{r}
ggplot(applications, aes(x = gender, y = processing_time, fill = race)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Paired") +  # You can choose other palettes like "Dark2", "Set3", etc.
  labs(title = "Distribution of Processing Time by Gender and Race",
       x = "Gender",
       y = "Processing Time (days)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Adjust text angle if needed
```



```{r}
# Create the boxplot without displaying outliers and limit y-axis to 50
ggplot(applications, aes(x = as.factor(experience_years), y = in_degree, color = experience)) +
  geom_boxplot(outlier.shape = NA) +  # This hides outliers
  labs(title = "Boxplot of In-Degree by Experience", x = "Experience", y = "In-Degree") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 20))  # Limit y-axis to 0 to 50
```



```{r}
# Create the boxplot using 'race' for x-axis and 'experience_years' to differentiate within each boxplot
ggplot(applications, aes(x = as.factor(gender), y = in_degree, fill = as.factor(experience_years))) +
  geom_boxplot(outlier.shape = NA) +  # This hides outliers
  scale_fill_brewer(palette = "Paired") +  # Use a color palette that provides clear differentiation
  labs(title = "Boxplot of In-Degree by Genger and Experience Years",
       x = "Gender",
       y = "In-Degree") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 20)) +  # Limit y-axis to 0 to 20 for clarity
  theme(legend.position = "bottom")  # Adjust legend position to avoid clutter
```


```{r}
# Create the boxplot using 'race' for x-axis and 'experience_years' to differentiate within each boxplot
ggplot(applications, aes(x = as.factor(gender), y = out_degree, fill = as.factor(experience_years))) +
  geom_boxplot(outlier.shape = NA) +  # This hides outliers
  scale_fill_brewer(palette = "Paired") +  # Use a color palette that provides clear differentiation
  labs(title = "Boxplot of Out-Degree by Genger and Experience Years",
       x = "Gender",
       y = "Out-Degree") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 20)) +  # Limit y-axis to 0 to 20 for clarity
  theme(legend.position = "bottom")  # Adjust legend position to avoid clutter
```

**Mean Proc Time by workgroup and show distribution of means**
```{r}
wg_means <- applications %>%
  group_by(workgroup) %>%
  summarise(mean_proc_time=mean(processing_time, na.rm = TRUE),
            n = n()) %>% arrange(desc(mean_proc_time))

# sort it
ggplot(wg_means, aes(x = reorder(workgroup, mean_proc_time), y = mean_proc_time)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Mean Processing Time by Workgroup",
       x = "Workgroup",
       y = "Mean Processing Time (days)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



**Application Status is limited to 150 and 161**
```{r}
applications %>%
  group_by(gender) %>%
  summarise(mean_app_proc_time=mean(processing_time, na.rm = TRUE),
            n = n()) %>% arrange(desc(n))


applications %>%
  group_by(race) %>%
  summarise(mean_app_proc_time=mean(processing_time, na.rm = TRUE),
            n = n()) %>% arrange(desc(n))
```


| Code | Status                                               | Frequency | Percent |
|------|------------------------------------------------------|-----------|---------|
| 150  | Patented Case                                        | 4,288,187 | 46.5%   |
| 161  | Abandoned -- Failure to Respond to an Office Action  | 1,107,772 | 12.0%   |
| 250  | Patent Expired Due to NonPayment of Maintenance Fees Under 37 CFR 1.362 | 919,203 | 10.0%   |
| 159  | Provisional Application Expired                     | 848,536   | 9.2%    |
| 30   | Docketed New Case - Ready for Examination           | 350,347   | 3.8%    |
| 218  | RO PROCESSING COMPLETED-PLACED IN STORAGE           | 284,700   | 3.1%    |
| 566  | PCT - International Search Report Mailed to IB      | 234,837   | 2.5%    |
| 41   | Non Final Action Mailed                              | 168,103   | 1.8%    |
| 166  | Abandoned -- File-Wrapper-Continuation Parent Application | 165,589 | 1.8%    |
| 19   | Application Undergoing Preexam Processing            | 126,741   | 1.4%    |
| N/A  | Other Codes                                          | 709,623   | 7.7%    |
|      | Missing                                              | 27,532    | 0.3%    |
|      | Total                                                | 9,231,170 | 100.0%  |




#### Sequence diff vs processing_time
```{r}
# Generate scatter plot of sequence_diff vs processing_time
ggplot(applications, aes(x = sequence_diff, y = processing_time)) +
  geom_point(alpha = 0.5) +  # alpha for transparency if many points overlap
  labs(x = "Sequence Difference", y = "Processing Time",
       title = "Scatter Plot of Sequence Difference vs Processing Time") +
  theme_minimal()  # A clean theme for the plot
```



```{r}

ggplot(applications, aes(x = out_degree, y = processing_time)) +
  geom_point(alpha = 0.5) +  # alpha for transparency if many points overlap
  labs(x = "out_degree", y = "Processing Time",
       title = "Scatter Plot of out_deg vs Processing Time") +
  theme_minimal()  # A clean theme for the plot
```


```{r}

ggplot(applications, aes(x = out_degree, y = processing_time)) +
  geom_point(alpha = 0.5) +  # alpha for transparency if many points overlap
  labs(x = "in_degree", y = "Processing Time",
       title = "Scatter Plot of out_deg vs Processing Time") +
  theme_minimal()  # A clean theme for the plot
```

```{r}

ggplot(applications, aes(x = betweenness, y = processing_time)) +
  geom_point(alpha = 0.5) +  # alpha for transparency if many points overlap
  labs(x = "betweenness", y = "Processing Time",
       title = "Scatter Plot of between vs Processing Time") +
  theme_minimal()  # A clean theme for the plot
```


```{r}

ggplot(applications, aes(x = advice_count_final)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 100) +
  labs(title = "Distribution of Experience",
       x = "advice_counts",
       y = "Frequency")
```


### Adivce Network Visualization
```{r}
edges <- read.csv("/Users/kaz/Desktop/Final Project_org/Data/edges_sample.csv")
advice_network <- graph_from_data_frame(d = edges[, c("ego_examiner_id", "alter_examiner_id")], directed = TRUE)
```



```{r}
# value count gender in applications
table(applications$gender)

# convert the "" to NA
applications$gender[applications$gender == ""] <- "NA"
table(applications$gender)
```


```{r}
# Convert identifiers to character to avoid issues with factors
edges$ego_examiner_id <- as.character(edges$ego_examiner_id)
edges$alter_examiner_id <- as.character(edges$alter_examiner_id)
applications$examiner_id <- as.character(applications$examiner_id)
# Ensure applications have unique entries per examiner
applications_unique <- applications %>% distinct(examiner_id, .keep_all = TRUE)

### Create the Advice Network
advice_network <- graph_from_data_frame(d = edges[, c("ego_examiner_id", "alter_examiner_id")], directed = TRUE)

### Merge Node Attributes
# First, merge node attributes with the main network
node_attributes <- merge(data.frame(id = V(advice_network)$name), applications_unique, by.x = "id", by.y = "examiner_id", all.x = TRUE)
V(advice_network)$race <- node_attributes$race
V(advice_network)$gender <- node_attributes$gender

### Analyze the Largest Subgraph
# Identify the largest component
largest_subgraph_members <- components(advice_network)$membership == which.max(sizes(components(advice_network)))

# Create the induced subgraph for the largest component
largest_advice_network <- induced_subgraph(advice_network, which(largest_subgraph_members))

# Transfer and map attributes from the main network to the subgraph
V(largest_advice_network)$gender <- V(advice_network)$gender[V(largest_advice_network)]
V(largest_advice_network)$race <- V(advice_network)$race[V(largest_advice_network)]
```




```{r}


in_degree_centrality <- degree(largest_advice_network, mode = "in") # in-degree centrality
out_degree_centrality <- degree(largest_advice_network, mode = "out")

# Add out-degree centrality as a vertex attribute for visualization
# Calculate betweenness centrality for each node (examiner)
betweenness_centrality <- betweenness(largest_advice_network, directed = TRUE)

V(largest_advice_network)$out_degree <- out_degree_centrality
V(largest_advice_network)$in_degree <- in_degree_centrality
V(largest_advice_network)$betweenness <- betweenness_centrality

# Create a dataframe of centrality scores
centrality_scores <- data.frame(
  examiner_id = V(largest_advice_network)$name,
  in_degree = in_degree_centrality,
  out_degree = out_degree_centrality,
  betweenness = betweenness_centrality
)
```

```{r}
# Calculate means and standard deviations for in-degree and out-degree
in_mean <- mean(degree(largest_advice_network, mode = "in"))
in_sd <- sd(degree(largest_advice_network, mode = "in"))
out_mean <- mean(degree(largest_advice_network, mode = "out"))
out_sd <- sd(degree(largest_advice_network, mode = "out"))
```

```{r}
gender_colors <- c(male = "slateblue", female = "orange", "NA" = "gray")

# Apply the color mapping
V(largest_advice_network)$color <- gender_colors[V(largest_advice_network)$gender]


png(filename = "Network_Visualization_Bet_by_Gender.png", width = 1600, height = 1600)

# Plot the network
plot(largest_advice_network,
     main = "Network Visualization, Size=Betweenness Centrality, Color=Gender",
     vertex.size = ifelse(betweenness(largest_advice_network, directed = TRUE) >
                                  quantile(betweenness(largest_advice_network, directed = TRUE), 0.75), 4, 1),
     vertex.label = NA,
     vertex.color = V(largest_advice_network)$color,  # Color by gender
     vertex.frame.color = NA,  # No outline for nodes
     edge.arrow.size = 0.1,
     edge.color = rgb(0.2, 0.2, 0.2, 0.1))  # Semi-transparent edges
dev.off()
```


```{r}
race_colors <- c(
  white = "#F0E442",  # Light Yellow, good for contrast with dark colors
  black = "#000000",  # Black, offers strong contrast against lighter colors
  Asian = "#FFA500",  # Orange, vibrant and stands out well
  Hispanic = "#377EB8",  # Blue, a different shade to distinguish from other colors
  other = "#E41A1C"   # Red, distinct and easily noticeable
)

V(largest_advice_network)$color <- race_colors[V(largest_advice_network)$race]


# Plot the network
png(filename = "Network_Visualization_Bet_by_Race.png", width = 800, height = 800)
plot(largest_advice_network,
     main = "Network Visualization by Race",
     vertex.size = ifelse(betweenness(largest_advice_network, directed = TRUE) >
                            quantile(betweenness(largest_advice_network, directed = TRUE), 0.75), 4, 1),
     vertex.label = NA,
     vertex.color = V(largest_advice_network)$color,  # Color by race
     vertex.frame.color = NA,  # No outline for nodes
     edge.arrow.size = 0.06,
     edge.color = rgb(0.2, 0.2, 0.2, 0.05))  # Semi-transparent edges
dev.off()
```


```{r}
gender_colors <- c(male = "slateblue", female = "orange", "NA" = "gray")

# Apply the color mapping
V(largest_advice_network)$color <- gender_colors[V(largest_advice_network)$gender]


png(filename = "Network_Visualization_by_Gender_indeg.png", width = 1600, height = 1600)

# Plot the network
plot(largest_advice_network,
     main = "Network Visualization, Size=In-degree Centrality, Color=Gender",
     vertex.size = ifelse(degree(largest_advice_network, mode = "in") >
                            quantile(degree(largest_advice_network, mode = "in"), 0.75), 4, 1),
     vertex.label = NA,
     vertex.color = V(largest_advice_network)$color,  # Color by gender
     vertex.frame.color = NA,  # No outline for nodes
     edge.arrow.size = 0.1,
     edge.color = rgb(0.2, 0.2, 0.2, 0.1))  # Semi-transparent edges
dev.off()
```


```{r}
race_colors <- c(
  white = "#F0E442",  # Light Yellow, good for contrast with dark colors
  black = "#000000",  # Black, offers strong contrast against lighter colors
  Asian = "#FFA500",  # Orange, vibrant and stands out well
  Hispanic = "#377EB8",  # Blue, a different shade to distinguish from other colors
  other = "#E41A1C"   # Red, distinct and easily noticeable
)

V(largest_advice_network)$color <- race_colors[V(largest_advice_network)$race]


# Plot the network
png(filename = "Network_Visualization_by_Race_indeg.png", width = 800, height = 800)
plot(largest_advice_network,
     main = "Network Visualization, Size=In-degree Centrality, Color=Race",
     vertex.size = ifelse(degree(largest_advice_network, mode = "in") >
                            quantile(degree(largest_advice_network, mode = "in"), 0.75), 4, 1),
     vertex.label = NA,
     vertex.color = V(largest_advice_network)$color,  # Color by race
     vertex.frame.color = NA,  # No outline for nodes
     edge.arrow.size = 0.06,
     edge.color = rgb(0.2, 0.2, 0.2, 0.05))  # Semi-transparent edges
dev.off()
```




